\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-11-13 redaction/dhp_july2017.tex
Changement rédaction

\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars at University of California-Irvine and Paris-Dauphine for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \& Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

@@

\date{October 2017}
 \maketitle
\bigskip

@@

Our three main findings can be summarized as follows.

First, we show that additive costs are quantitatively sizeable. On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model. To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}. This represents a valuable insight for calibrating related international trade or business cycle models.

@@

Obviously, this dataset is not immune of limitations. First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs. It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}. However, it is also clear that transport costs do represent a sizeable share of trade costs as pointed out in Section \ref{sec:Intro}. Further, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit. In this respect, this dataset embraces only a partial view of international transport barriers. Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e. ignoring those flows that did not happen because of presumably prohibitive transport costs. In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs. %JH: OK, parfaitement clair pour moi !

Yet, this dataset also has (at least) three advantages. First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source. Based on customs declarations, the US Imports database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system. This will be crucial to compute transport costs, as we detail below. In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air. We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode. Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country). This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs. Third, this dataset is available over a long time span. We are thus able to exploit information on a yearly basis from 1974 to 2013. \cite{hummels2007} preceded us in investigating the time dimension of this database. However, our paper differs from his work in two main respects. First, we extend the time coverage up to 2013, while \cite{hummels2007} stops in 2004. Second, and most importantly, we study the time trends of transport costs by disentangling the additive from the iceberg components. This proves to be key in the characterization of transport costs changes over time, in particular regarding the role of the composition effects.\smallskip

@@

Our first contribution to the literature is to provide estimates for the size of both the ad-valorem and the additive components of transport costs. Table \ref{tab:summary_results} reports a summary of our results. It displays the mean and median values of each type of trade costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data. The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})). Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level. Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year. Overall results (over the whole period) are not substantially affected if this year is included though. This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}

@@

Table \ref{tab:summary_results} calls for two types of comments. The first type is of statistical order. As displayed in Table \ref{tab:summary_results}, mean values are systematically higher than medians, by around one percentage point. This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median. One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point. This is first due to the fact that observed data are \emph{de facto} trade-weighted (i.e., by the weight of each good among total trade), while our estimates are based on simple arithmetic means - Section \ref{sec:results_trends} will go into more details of these compositions effects. Furthermore, Table \ref{tab:summary_results} reports values/percentages, whereas all our estimates are based on log-linearization. Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values. Descriptive statistics showing that it is indeed the case are available upon request to the authors.\medskip

@@

\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year. As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, ee estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other). Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
\end{eqnarray}
As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)). Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions. Both equations are estimated by transport mode.

@@

Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded). Details are reported in Appendix \ref{app:comp-effects}).

@@

\section{Time trends in transport costs: Robustness to the weighting scheme}

@@

The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives tow comments. First, differences do show up, in particular for the multiplicative component of Air transport. This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise. However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than to partially offset as argued by \cite{hummels2007}. Rather than the weighing scheme, this confirms that the functional form and precisely the modeling of the additive component, turns out to be key in understanding the underlying determinants of the overall transport costs time trends, as we pointed out in Section \ref{sec:results_trends}.

>>>>

\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \& Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

@@

\date{November 2017}
 \maketitle
\bigskip

@@

Our three main findings can be summarized as follows. First, we show that additive costs are quantitatively sizeable. On average over 1974-2013, the additive cost is estimated to 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport model. To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs.\footnote{Given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}. This represents a valuable insight for calibrating related international trade or business cycle models.

@@

Obviously, this dataset is not immune of limitations. First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs. It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}. Further, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit. In this respect, this dataset embraces only a partial view of international transport barriers. However, it is also clear that direct transport costs do represent a sizeable share of trade costs as pointed out in Section \ref{sec:Intro}. Last, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e. ignoring those flows that did not happen because of presumably prohibitive transport costs. In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs. %JH: OK, parfaitement clair pour moi !

Yet, this dataset also has (at least) three advantages. First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source. Based on customs declarations, the US Imports database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system. This will be crucial to compute transport costs, as we detail below. In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air. We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode. Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country). This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs. Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013. \smallskip

@@

Our first contribution to the literature is to provide estimates for the size of both the ad-valorem and the additive components of transport costs. Table \ref{tab:summary_results} reports a summary of our results. It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data. The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})). Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level. Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year. Overall results (over the whole period) are not substantially affected if this year is included though. This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}

@@

Table \ref{tab:summary_results} calls for two types of comments. The first type is of statistical order. As displayed in Table \ref{tab:summary_results}, mean values are systematically higher than medians, by around one percentage point. This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean up compared to the median. One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point. The key explanation for this is of statistical order. Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization. Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values. Descriptive statistics showing that that both observed and predicted values do match in logarithms are available upon request to the authors.\medskip

@@

\noindent where $\mathbb{1}_i$ and $\mathbb{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year. As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other). Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}
%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbb{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
\end{eqnarray}
As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three components: The country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)). Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions. Both equations are estimated by transport mode.

@@

Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded). Details are reported in Appendix \ref{app:comp-effects}.

@@

\section{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}

@@

The comparison of Figure \ref{fig:compeffects_robustness} (applying Hummels's (2007) weighting scheme) and Figure \ref{fig:totalTC_compeffects_excl} (applying our weighing scheme) drives two comments. First, differences do show up, in particular for the multiplicative component of Air transport. This suggests that the weighting scheme is not innocuous in the time trend decomposition exercise. However, Figure \ref{fig:compeffects_robustness} also shows that the composition effects have contributed to strengthen the decrease in the ``ceteris paribus'' transport costs in Air transport (in Figure \ref{fig:compeffects_robustness}, panel (a), overall transport costs have decreased more that the fitted component), in accordance with the conclusions drawn from Figure \ref{fig:totalTC_compeffects_excl}, rather than to partially offset as argued by \cite{hummels2007}. Rather than the weighing scheme, this confirms that the functional form and precisely the modeling of the additive component, turns out to be key in understanding the underlying determinants of the overall transport costs time trends, as we pointed out in Section \ref{sec:results_trends}.
